---
title: "Beer Case Study Report"
author: "Cho Kim, Nicole Witltin, and Jonathan Marin"
date: "February 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load libraries
library(sqldf)
library(magrittr)
library(kableExtra)
library(ggplot2)
library(maps)
library(knitr)
```

## Introduction

Introduction to the project. The introduction should not reference a project, persay. No
part of this should be informal.

The introduction needs to be written as if you are presenting the work to someone who
has given you the data to analyze and wants to understand the result. Pretend its a
presentation for a client. This may take some imagination of whom your client might be.
If it sounds like a student presentation, that is not acceptable.


#### Breweries present in each state

```{r echo=TRUE}
#Read the data
Beers <- read.csv("Beers.csv", encoding='UTF-8')
Brews <- read.csv("Breweries.csv", encoding='UTF-8', sep=",")

#calculate the number of breweries per state
State_Brews <-  aggregate(Brew_ID ~ State , Brews, length)

colnames(State_Brews) <- c("State","NumberofBreweries")

kable(State_Brews, "html") %>% kable_styling(bootstrap_options="striped", full_width = F, font_size = 14)

```

#### First 6 observations 

```{r echo = TRUE}

Beer_Data <- sqldf("select a.Brew_ID, a.Name as 'BrewName', a.City, a.State, b.Name as 'BeerName', b.ABV, b.IBU, b.Style, b.Ounces from Brews a join Beers b on a.Brew_ID = b.Brewery_id")

kable(head(Beer_Data), "html") %>% kable_styling(bootstrap_options="striped", full_width = F, font_size = 14)

```

#### Last 6 observations 

```{r echo = TRUE}

kable(tail(Beer_Data), "html") %>% kable_styling(bootstrap_options="striped", full_width = F, font_size = 14)

```

#### Number of NA's for each column

```{r echo = TRUE}

Brew_ID_NA <-sum(is.na(Beer_Data$Brew_ID))
BrewName_NA<-sum(is.na(Beer_Data$BrewName))
City_NA <- sum(is.na(Beer_Data$City))
State_NA <- sum(is.na(Beer_Data$State))
BeerName_NA <- sum(is.na(Beer_Data$BeerName))
ABV_NA <- sum(is.na(Beer_Data$ABV))
IBU_NA <- sum(is.na(Beer_Data$IBU))
Style_NA <- sum(is.na(Beer_Data$Style))
Ounces_NA <- sum(is.na(Beer_Data$Ounces))

na_columns <- data.frame(matrix(ncol =2, nrow=0))

colnames(na_columns) <- c("ColumnName", "NumberOfNAs")

na_columns[nrow(na_columns) +1,] = c("Brew_ID", Brew_ID_NA)
na_columns[nrow(na_columns) +1,] = c("BrewName", BrewName_NA)
na_columns[nrow(na_columns) +1,] = c("City", City_NA)
na_columns[nrow(na_columns) +1,] = c("State", State_NA)
na_columns[nrow(na_columns) +1,] = c("BeerName", BeerName_NA)
na_columns[nrow(na_columns) +1,] = c("ABV", ABV_NA)
na_columns[nrow(na_columns) +1,] = c("IBU", IBU_NA)
na_columns[nrow(na_columns) +1,] = c("Style", Style_NA)
na_columns[nrow(na_columns) +1,] = c("Ounces", Ounces_NA)


kable((na_columns), "html") %>% kable_styling(bootstrap_options="striped", full_width = F, font_size = 14)

```

#### Median alcohol content (ABV) and International Bitterness Unit (IBU) for each state
```{r echo = TRUE}

ABV_Median <- aggregate(ABV ~ State, data = Beer_Data, median, na.action = na.omit)
IBU_Median <- aggregate(IBU ~ State, data = Beer_Data, median, na.action = na.omit)

median_table <- sqldf("select a.State, ABV as 'ABV_Median', b.IBU as 'IBU_Median' from ABV_Median a left join IBU_Median b on a.State = b.State")

#calculate percent for each median ABV
median_table <- within(median_table, ABVinPercent <- ABV_Median*100)
#omit ABV_Median column for display
median_table$ABV_Median <- NULL
#reorder columns so ABV is first
median_table <- median_table[c(1,3,2)]
kable((median_table), "html") %>% kable_styling(bootstrap_options="striped", full_width = F, font_size = 14)

```

#### Bar chart for Median Alcohol content and International Bitterness Unit (IBU)
```{r echo = TRUE, fig.width=12}
median_table$State <- factor(median_table$State, levels= median_table$State[order(-median_table$ABVinPercent)])
g <- ggplot(median_table, aes(x=State, y=ABVinPercent))
g +geom_bar(stat="identity") + labs(title = "Median ABV(%) by State", y="ABV in %")

median_table$State <- factor(median_table$State, levels= median_table$State[order(-median_table$IBU_Median)])
g <- ggplot(median_table, aes(x=State, y=IBU_Median))
g +geom_bar(stat="identity") + labs(title = "Median IBU by State", y="IBU")
```

#### Maximum alcohol content (ABV) and International Bitterness Unit (IBU) for each state
```{r echo = TRUE}
#calculate % for ABV
Beer_Data <- within(Beer_Data, ABVinPercent <- ABV*100)

maxABV <- which.max(Beer_Data$ABV)

kable((Beer_Data[maxABV,]), "html") %>% kable_styling(bootstrap_options="striped", full_width = F, font_size = 14)

Beer_Data$State[maxABV]

maxIBU <- which.max(Beer_Data$IBU)

kable((Beer_Data[maxIBU,]), "html") %>% kable_styling(bootstrap_options="striped", full_width = F, font_size = 14)

Beer_Data$State[maxIBU]

```

The state with the maximum ABV is Colorado. The beer style is a Quadrupel. It is a Belgian Style Quadrupel ale that has 12.8% ABV. 

The state with the most bitter beer (in IBUs) is Oregon. The beer style is an American Double / Imperial IPA. 

#### Summary statistics for the ABV Variable
```{r echo = TRUE}
summary_ABV <- (summary(Beer_Data$ABVinPercent))
summary_ABV
```

#### ABV vs IBU scatter plot
```{r echo = TRUE}
ggplot(Beer_Data, aes(x=IBU, y=ABVinPercent)) + geom_point()+labs(y="ABV(%)", title = "ABV in % vs IBU")
```